Place of death in NRS COVID-19 deaths

### 2021-03-08 covid_nrs_place_of_death.Rmd
changed it so that we don't re-download shapefiles every time;
fixed order of .rmd script;

added a cumulative deaths graph, needs more work to be clear

@@

## Number of deaths by age group

```{r, fig.width=12, fig.height=9, warning=FALSE}
## graph format copied from: https://github.com/davidhen/nrs_covid_stats/
weekly_deaths_overall %>% 
  filter(sex == "all" & place_of_death == "All") %>% 
  filter(age != "all") %>%
  group_by(age) %>% 
  summarise(n = sum(deaths_all_causes)) %>%
  mutate(pct = round(n/sum(n), 3)) %>% 
  ggplot(aes(age, n)) +
  geom_col(fill = "#4477AA") +
  geom_text(aes(label = scales::percent(pct, accuracy = 0.1), y = n + 1000)) +
  # theme_minimal() +
  labs(x = "",
       y = "",
       title = "Deaths Associated with COVID-19 by Age Group",
       subtitle = date_range_text,
       caption = paste0(c(date_range_text,source_nrs_text), collapse="\n")
  )
```

@@

figure_annual_deaths_by_age_group_2020 <-
  total_deaths_age %>% 
  filter(year==2020) %>%
  filter(place_of_death != "Other") %>%
  ggplot(aes(x = age, y = deaths_all_causes)) +
  # geom_col(fill = "#4477AA", alpha = 0.5, colour = "grey10") +
  geom_col(fill = "grey20", colour = "grey10", alpha = 0.5) +
  geom_col(aes(x=age, y=annual_deaths_max), fill = "#4477AA", colour = "grey10", alpha = 0.5) +  # historical maximum over the actual figure
  geom_point(aes(x=age, y=annual_deaths_mean)) +
  geom_errorbar(aes(x = age, ymin = annual_deaths_min, ymax = annual_deaths_max)) +
  facet_wrap(~place_of_death) +
  scale_y_continuous(labels = function(x) format(x,big.mark=",")) +
  # theme_minimal() +
  labs(x = "Age",
       y = "Deaths (N)",
       title = "Deaths in 2020 by age group & place of death",
       subtitle = paste0(c(only_2020_text,light_shaded_area_shows_difference),collapse="\n"),
       caption = paste0(c(excluded_other_place_2020,bar_and_error_bars_2020_text,death_registrations_text,source_nrs_text), collapse="\n")
  )

>>>>

# Cumulative deaths

Note: I've left this unfinished, it doesn't tell a better story than the weekly figures!

```{r}
text_non_covid_deaths_equivalent_to_all_cause_before_wk12 <- "Note: all-cause deaths and non-covid deaths were considered equivalent before 2020, week 12"

figure_cumulative_deaths_by_place <-
  merged_deaths_overall %>%
  filter(sex == "all" & age == "all") %>%
  filter(!place_of_death %in% c("All","Other")) %>%
  mutate(place_of_death = factor(place_of_death, levels = c("Hospital", "Home & other non-institution", "Care home"))) %>%  # needs to be factor with same levels, otherwise it messes up level order when added as geom_text later!)
  mutate(deaths_non_covid = if_else(condition = is.na(deaths_non_covid), true = deaths_all_causes, false = deaths_non_covid)) %>%  # replace NA in non-covid deaths with all cause deaths - before covid, non-covid deaths are equivalent to all deaths 
  pivot_longer(cols = c(deaths_all_causes, deaths_non_covid), 
               names_to = "deaths_type", values_to = "deaths") %>%
  mutate(deaths_type = case_when(
    deaths_type == "deaths_all_causes" ~ "All-cause deaths",
    deaths_type == "deaths_non_covid" ~ "Non-COVID deaths",
    TRUE ~ NA_character_
  )) %>%
  group_by(place_of_death, deaths_type) %>%
  mutate(
    cumulative_deaths = cumsum(replace_na(deaths, replace = 0)),
    cumulative_historical_mean = cumsum(replace_na(deaths_mean, replace = 0)),
    cumulative_historical_min = cumsum(replace_na(deaths_min, replace = 0)),
    cumulative_historical_max = cumsum(replace_na(deaths_max, replace = 0))
    ) %>%
  ungroup %>%
  ggplot(aes(x = date_w_c)) +
  scale_x_date(date_labels = "%d-%m-%y", date_breaks = "8 weeks", limits = c(earliest_date, most_recent_date_available_2021)) +
  scale_y_continuous(labels = function(x) format(x, big.mark = ",")) +
  geom_ribbon(aes(ymin = cumulative_historical_min, ymax = cumulative_historical_max, fill = place_of_death), alpha = 0.4) +
  geom_line(aes(y = cumulative_historical_mean, colour = place_of_death), linetype = "dashed", alpha = 0.8, show.legend = FALSE) +
  geom_line(aes(y = cumulative_deaths, colour = place_of_death)) +
  facet_wrap(~deaths_type) +
  scale_colour_viridis_d(option = "C") +
  scale_fill_viridis_d(option = "C") +
  theme(
    legend.position = "top",
    plot.caption = element_text(size = 10, colour = "gray60"),
    plot.margin = margin(5, 15, 5, 5),
    axis.text.x = element_text(angle = 60, colour = "gray40", vjust=1, hjust=1)
  ) +
  labs(x = "Date w/c",
       y = "Cumulative deaths (N)",
       title = "Cumulative deaths by place of death",
       # subtitle = paste0(c(only_2020_text,light_shaded_area_shows_difference),collapse="\n"),
       caption = paste0(c(text_non_covid_deaths_equivalent_to_all_cause_before_wk12,death_registrations_text,source_nrs_text), collapse="\n")
  ) +
  NULL

save_output_file(filename = paste0(dir_outputs,"/figure_cumulative_deaths_by_place"), extensions = c(".pdf",".svg",".png"), plot = figure_cumulative_deaths_by_place, width = 15, height = 12, dpi = 300, units = "cm")

figure_cumulative_deaths_by_place
```

@@

@@

figure_annual_deaths_by_age_group_2020 <-
  total_deaths_age %>% 
  filter(year==2020) %>%
  filter(place_of_death != "Other") %>%
  ggplot(aes(x = age, y = deaths_all_causes)) +
  # geom_col(fill = "#4477AA", alpha = 0.5, colour = "grey10") +
  geom_col(fill = "grey20", colour = "grey10", alpha = 0.5) +
  geom_col(aes(x=age, y=annual_deaths_max), fill = "#4477AA", colour = "grey10", alpha = 0.5) +  # historical maximum over the actual figure
  geom_point(aes(x=age, y=annual_deaths_mean)) +
  geom_errorbar(aes(x = age, ymin = annual_deaths_min, ymax = annual_deaths_max)) +
  facet_wrap(~place_of_death) +
  scale_y_continuous(labels = function(x) format(x,big.mark=",")) +
  labs(x = "Age",
       y = "Deaths (N)",
       title = "Deaths in 2020 by age group & place of death",
       subtitle = paste0(c(only_2020_text,light_shaded_area_shows_difference),collapse="\n"),
       caption = paste0(c(excluded_other_place_2020,bar_and_error_bars_2020_text,death_registrations_text,source_nrs_text), collapse="\n")
  )
